---
title: "Capstone Exercise Anderson"
author: "Sarah Anderson"
date: "Tuesday, April 28, 2015"
output: html_document
---

##Capstone Exercise for the Software Carpentry Workshop

We will be using the Rmd files to create a short analysis report of the gapminder dataset.  

The gapminder dataset includes data on GDP, life expectancy, and population for various years for each country worldwide.

```{r, LoadingData, echo=FALSE, message=FALSE}
gapminder<- read.csv("data/gapminder.csv")
head(gapminder)
```

Three countries of interest are Brazil, Tanzania, and France. We are interested in how GDP has changed in each of these three countries overtime.  
  
  

```{r, LoadingLibraries, echo=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
```

###Brazil
```{r, BrazilPlot, echo=FALSE, message=FALSE, fig.height=3}
gapminder %>%
  filter(country == "Brazil") %>%
  ggplot(aes(x=year, y=gdpPercap)) + geom_point(color="darkblue", size = 3) + geom_line(color="violet")
```
  
###Tanzania
```{r, TanzaniaPlot, echo=FALSE, message=FALSE, fig.height=3}
gapminder %>%
  filter(country == "Tanzania") %>%
  ggplot(aes(x=year, y=gdpPercap)) + geom_point(color="darkblue", size = 3) + geom_line(color="violet")
```


###France
```{r, FrancePlot, echo=FALSE, message=FALSE, fig.height=3}
gapminder %>%
  filter(country == "France") %>%
  ggplot(aes(x=year, y=gdpPercap)) + geom_point(color="darkblue", size = 3) + geom_line(color="violet")
```

All three countries have experience growth over the last 60 years, but each of these countries have experience a different pattern. Brazil has experience periods of rapid growth with some periods of _slight decline but a consistent trend in growing GDP_. Tanzania experienced rapid growth in the 1960s and 1970s but then had a long period of _decline before exponential growth_ starting around 2000. France has experienced _consistent linear growth_ throughout the whole time period. 

##Life Expectancy by Continent for 2007
```{r, LifeExpTable, echo=FALSE, message=FALSE, fig.height=3}
gapminder %>%
  group_by(continent) %>%
  summarize(Mean=mean(lifeExp), Minimum=min(lifeExp), Maximum=max(lifeExp))
```
  
Oceania and Europe are the regions with _highest_ average life expectancy while Africa is the region with the _lowest_ average life expectancy.  Africa and Asia are regions with countries that have some of the _lowest_ life expectancies while all regions have some countries with relatively _high_ life expectancy. 